Learning CNN filters from user-drawn image markers for coconut-tree image classification

### 2020-07-24 IEEE-GRSL-Response-Letter.tex

\rwr{2} Pg. 5 - Line 6: An error in the paper title is something that should be avoided at all costs as it is very confusing for the reader.  "...User-drawn Image MAKERS..." ??? The first word of the introduction is also wrong: "DePP learning"? Overall, there are not many language errors, but a proof-reading is highly recommended

\ans Thanks. We have fixed them and revised the entire text to avoid typos. \eans

\rwr{2} Pg. 9 - Lines 14-18: Justify the choice of all hyperparameters (E.g. epochs, batch size, learning rate, etc). Have any previous experiments been made to tune these hyperparameters? Has the same dataset been used?

\ans Thanks to have pointed it out. We are now explaining that the dataset is randomly divided into training (200 images), validation (1000 images), and testing (remaining images) sets. We repeated each experiment three times, each using one of these random splits. However, the validation set was used only once to choose the hyperparameters of the network. We have not developed/used any optimization technique for this purpose. We have just executed tests with a few hyperparameter choices in order to select one for the experiments. The idea was to select the simplest and yet effective network in order to demonstrate that FLIM can create competitive solutions. We are clarifying this issue (Section III) and indicating it for future work (Section IV). 
\eans

@@

\rwr{2} Pg. 10 - Table 1: The F-score difference from FLIM+MLP (0.854) and VGG+FT (0.851) is statistically significant considering the standard deviation?  Has a statistical hypothesis test been applied?

For three splits only and 200 training images, it was not possible to claim that this difference is statistically significant. However, they are undoubtedly competitive, being FLIM+MLP a much simpler (lighter) network than VGG-FT. Even with a new experiment in which we increase the training set up to 1200 images, the difference increases in favor of FLIM+MLP, but they are still not statistically significant (except for one case with 1100 images). On the other hand, the f-score curve of FLIM+MLP was consistently above one of VGG-FT. We have improved the experimental section to clarify this and several other issues.  
\eans

@@

For three splits only and 200 training images, it was not possible to claim that this difference is statistically significant. However, they are undoubtedly competitive, being FLIM+MLP a much simpler (lighter) network than VGG-FT. Even with a new experiment in which we increase the training set up to 1200 images, the difference increases in favor of FLIM+MLP, but they are still not statistically significant (except for one case with 1100 images). On the other hand, the f-score curve of FLIM+MLP was consistently above one of VGG-FT. We have improved the experimental section to clarify this and several other issues.  
\eans

\rwr{2} Pg. 11 - It would be worth investigating or discussing if the time taken to mark some images is always less than the time take to produce more labeled images (E.g.: how many new labeled images would take for VGG+FT beat FLIMP+MLP?). Is there a trade-off?

\ans Thanks a lot for raising these questions. We noticed that it was not clear that markers are drawn on a few images selected from the training set (we used only four images). The user takes a couple of minutes to select images and draw markers in those images. This is much less than the time taken to examine and label 200 images. We also added an experiment with increasing training set sizes. VGG-FT could never beat FLIM+MLP. The reason is that the image features of FLIM are better than those of VGG-FE (the one extracted from the last convolutional layer of VGG-FT), and both use a similar MLP (the only difference is its input layer). So, higher is the number of training samples, MLP can improve the f-score of both, being the one of FLIM+MLP always above the one o

### 2020-07-24 main.tex
Updates from Overleaf

To reduce the need for large annotated training sets by exploiting the user knowledge, we present a method, called \textcolor{red}{\emph{Feature Learning from Image Markers} (FLIM)}, that needs a minimal set of images to learn the filters of each convolutional layer. The user selects the number of convolutional layers, their filter sizes, \textcolor{red}{a few images from a training set}, and draws markers in image regions that best discriminate the classes. \textcolor{red}{By clustering patches extracted from marker pixels at the input of each convolutional layer, FLIM finds filters that enhance those regions at the convolutional layers' output. The remaining training images can then be used to train the fully connected layers or another classifier.} The CNN architecture could also be optimized based on visual analytics~\cite{SpinnerTVCG2020}, but we have not explored this option yet. \textcolor{red}{In interactive semantic segmentation, user-selected clicks can affect part of the network to improve segmentation, but the filters are not designed from those markers~\cite{SofiiukCVPR2020}.}

@@

\textcolor{red}{For these experiments, we have selected the following models that combine feature extraction and classification methods.
 \begin{itemize}
\item FLIM+SVM: Image classification is performed by support vector machines (SVM) from FLIM-based features.  
\item FLIM+MLP: Image classification is performed by multi-layer perceptron (MLP) from FLIM-based features.  
\item FLIM-FT+MLP: The architecture in FLIM+MLP is fine-tuned by backpropagation. 
\item CL+MLP: The architecture in FLIM+MLP is trained from scratch by backpropagation.
\item VGG: VGG-16~\cite{simonyan2014very} is trained from scratch.
\item VGG-FT: VGG-16 is pre-trained on ImageNet and fine-tuned on the coconut-tree training set.
\item VGG-FE+SVM: Image features are obtained from the last convolutional layer of VGG-FT.
\end{itemize}
}

@@

\textcolor{red}{To evaluate H3, we used the validation set to increase the classifiers' number of training images. Note that the marker selection for FLIM was still the same from those four images. To our surprise, as the training set increases, the mean differences in f-score, as obtained among the three splits, favor FLIM+MLP over FLIM-FT+MLP (H2), VGG-FT (H1 and H3), and CL+MLP (H2). The differences between FLIM+MLP and CL+MLP demonstrate that FLIM is superior to backpropagation in filter estimation (H2). Finally, Table II shows the mean results of FLIM+SVM among the three splits as the number of images for marker selection increases, and the training set size remains with 200 images (H4). The results are not conclusive for H4, but the choice of more markers does not deteriorate the performance of FLIM. This point deserves further investigation -- i.e., choosing relevant training images and markers for FLIM. }

>>>>

To reduce the need for large annotated training sets by exploiting the user knowledge, we present a method, called \textcolor{red}{\emph{Feature Learning from Image Markers} (FLIM)}, that needs a minimal set of images to learn the filters of each convolutional layer. The user selects the number of convolutional layers, their filter sizes, \textcolor{red}{a few images from a training set}, and draws markers in image regions that best discriminate the classes. The user takes a couple of minutes to select images and draw markers in those images. \textcolor{red}{By clustering patches extracted from marker pixels at the input of each convolutional layer, FLIM finds filters that enhance those regions at the convolutional layers' output. The remaining training images can then be used to train the fully connected layers or another classifier.} The CNN architecture could also be optimized based on visual analytics~\cite{SpinnerTVCG2020}, but we have not explored this option yet. \textcolor{red}{In interactive semantic segmentation, user-selected clicks can affect part of the network to improve segmentation, but the filters are not designed from those markers~\cite{SofiiukCVPR2020}.}

@@

\textcolor{red}{For these experiments, we have selected the following models that combine feature extraction and classification methods.
 \begin{itemize}
\item FLIM+SVM: Image classification is performed by support vector machines (SVM) from FLIM-based features.  
\item FLIM+MLP: Image classification is performed by \textcolor{red}{Multi-Layer Perceptron} (MLP) from FLIM-based features.  
\item FLIM-FT+MLP: The architecture in FLIM+MLP is fine-tuned by backpropagation. 
\item CL+MLP: The architecture in FLIM+MLP is trained from scratch by backpropagation.
\item VGG: VGG-16~\cite{simonyan2014very} is trained from scratch.
\item VGG-FT: VGG-16 is pre-trained on ImageNet and fine-tuned on the coconut-tree training set.
\item VGG-FE+SVM: Image features are obtained from the last convolutional layer of VGG-FT.
\end{itemize}
}

@@

\textcolor{red}{To evaluate H3, we used the validation set to increase the classifiers' number of training images. Note that the marker selection for FLIM was still the same from those four images. To our surprise, as the training set increases, the differences in mean f-score, as obtained among the three splits, favor FLIM+MLP over FLIM-FT+MLP (H2), VGG-FT (H1 and H3), and CL+MLP (H2). The differences between FLIM+MLP and CL+MLP demonstrate that FLIM is superior to backpropagation in filter estimation (H2). Finally, Table II shows the mean results of FLIM+SVM among the three splits as the number of images for marker selection increases, and the training set size remains with 200 images (H4). The results are not conclusive for H4, but the choice of more markers does not deteriorate the performance of FLIM. This point deserves further investigation -- i.e., choosing relevant training images and markers for FLIM. }
